---
title: "NBA Hall of Fame Prediction"
author: "Justin Mai"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
    keep_tex: true
fontsize: 11pt
documentclass: article
mainfont: "Times New Roman"
geometry: margin=1in
---

```{r include=FALSE}
library(tidyverse)
library(corrplot)
```


\newpage

# Abstract

\newpage

# Introduction

The NBA Hall of Fame inducts the most influential players, coaches, teams, and referees
yearly. There have only been just over 150 NBA players inducted to the Hall of Fame which
started with the inaugural class of 1959. So what classifies an NBA player as a Hall of Famer
compared to other NBA players? This research paper will identify the probability of current NBA players
one day making the Hall of Fame based on historical trends.

## Accolades and Awards

A common debate within the basketball community is the infamous "LeBron vs. Michael Jordan" debate
to crown on player as the Greatest of All Time (G.O.A.T.). Analysts will often start with the quantitative
in game statistics by looking at all time averages for both players. Looking at the primary statistics, throughout
Jordan's career, he averaged $30.1$ points per game, $6.2$ rebounds per game, and $5.3$ assists per game. On the other
end, LeBron averages $27.0$ points per game, $7.5$ rebounds per game, and $7.4$ assists per game. While Jordan has the
edge on scoring, LeBron has the edge in the other primary statistics so its difficult to make a clear conclusion
based on this. However, this isn't the primary argument for both players, if you've ever been part of this debate
you'll often hear the notion that "Jordan went 6 for 6 in championship games". The accolades and awards that each
player compiles is often the primary argument.

While there is no clear calculator for identifying if a player will make the Hall of Fame, in all cases, accolades and
awards will be a significant predictor to identifying HOFers. These accolades will consist of **Regular Season MVPs, Championship Wins, Finals MVPs, All-NBA Selections, All-Star Selections, End-of-Season Awards** and possibly much more. These awards all signify the impact
that a player has had on their respective teams, demonstrating how their contribution leads to the team's success.

## Player Impact / Other Considerations

Within the G.O.A.T debate, a primary argument for LeBron would be his longevity and consistent impact on the game
and the teams he goes to, the qualitative factors that goes beyond the box score and award counts. While accolades and awards 
provide a quantitative summary of a player's career, qualitative aspects such as leadership, clutch performances, career longevity, 
influence on team culture, and global popularity often shape the broader legacy of a player.

For example, LeBron’s ability to lead multiple franchises to the NBA Finals—winning championships with three different teams—is a testament 
to his versatility and value as a player. Similarly, players like Allen Iverson and Vince Carter are celebrated not only for their 
statistics and accolades, but also for their cultural impact, influence on future generations, and overall contribution to the evolution of the game.

When making predictions for Hall of Fame inductees, there are also many traits outside of the box score that contributes to
the selection process. However, this type of impact often correlates with the accolades and overall quantitative statistics so
it will be contributing factor within the analysis. Something like longevity will also be taken into account through the number
of years they played all together and the number of years they played for one team.

# Methods

```{r}
df <- read.csv("data/final.csv")
```

```{r}
train <- read.csv("data/train.csv")
test <- read.csv("data/test.csv")
```

```{r}
M = glm(hof ~ num_all_star + num_season + pts_per_game + dpoy, family = binomial(link="logit"), data = train)
M1 = glm(hof ~ num_all_star + pts_per_game + dpoy, family = binomial(link="logit"), data = train)

anova(M, M1)
```



# Results

# Discussion

## Limitations

## Next Steps

# Bibliography

https://www.sportingnews.com/us/nba/news/michael-jordan-vs-lebron-james-goat-debate/sl8xdozy5u1m1s4t5m3npeqo1

https://www.kaggle.com/datasets/sumitrodatta/nba-aba-baa-stats?select=Player+Career+Info.csv
